---
title: "AllData"
author: "MPG"
date: "8/24/2020"
output: html_document
---

1) Package load/retrieval
```{r,warning=F}
#necessary packages: 
pack_list<-list ("ggplot2","openair","dygraphs","xts","dataRetrieval","dplyr","plyr","tidyr","forecast","lubridate", "expss","gganimate","mgcv","mvtnorm","statmod","janitor")
for(i in 1:length(pack_list)){
  p <- pack_list[[i]]
  install.packages(p,repos = "https://cran.rstudio.com",dependencies = TRUE)
}

library(ggplot2)
library(dplyr)  
library(lubridate)
library(xts)
library(tidyverse)
library(dygraphs)
library(forecast)
library(timeSeries)
library(glue)
library(expss)
detach("package:dplyr", unload=TRUE) : library(dplyr)
options(stringsAsFactors = FALSE)

setwd("R:/EcosystemEcologyLab/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/01_input")

#######################################
```

2) LOOPED SITE ANALYSIS---saves a file with all sites and min,max, mean, and diel range of temperature data into EnviroDIY/AllData
```{r setup, include=FALSE}

### Code selects each file and summarize by 4 paramaters (min,max,mean, and diel range)
diel <- function(x){
  max(x) - min(x)} #create a function to calculate diel change 

file.names <- dir(getwd(), pattern =".csv") #a vector with all file names in the inout folder
temp.summary<-data.frame(matrix(ncol = 8, nrow = 0)) #empty dataframe that will contain the integrated files into one single dataframe
colnames(temp.summary) <- c("year","month","day","min","mean","max","diel","site") #name for the eight columns

         
for(i in 1:length(file.names)){  
              
        file <- read.table(file.names[i],header=TRUE, sep=",", stringsAsFactors=TRUE)
        data2 <- select(file,c(sitecode,dateEST,TimeEST,depth_mm,temp_degc))
        data2$dateEST <- as.Date(data2$dateEST, format = "%d %B %y")
        data2$dateEST <- as.character(data2$dateEST) ; data2$TimeEST <- as.character(data2$TimeEST)
        data2$datetime <- paste(data2$dateEST, data2$TimeEST) ; data2$datetime <- ymd_hm(data2$datetime)
        temp_data<-select(data2,c(sitecode,depth_mm,temp_degc,datetime))

              temp_data <- temp_data %>%
                mutate(year = year(datetime))  %>%
                mutate(month = month(datetime))  %>%
                mutate(day = day(datetime))
              temp_data <- temp_data %>%
                filter(year==c("2018"))
              temp_diel_data <- temp_data %>%
                group_by(year,month,day) %>%
                summarise_at("temp_degc", tibble::lst(min, mean, max, diel))

                        temp_diel_data$site <- as.vector(rep(file.names[i], nrow(temp_diel_data)))
                        colnames(temp_diel_data) <- c("year","month","day","min","mean","max","diel","site")
                        temp_diel_data<-as.data.frame(temp_diel_data)
                        temp.summary<-rbind(temp.summary,temp_diel_data)
                            }
                            
write.csv(temp.summary,file = "R:/EcosystemEcologyLab/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/04_product/AllData/AllDataSummarized.csv") #write the final file into the sensor product folder

```    

3) GENERAL PLOT---animation with monthly averages per site
```{r}
library(gganimate)
theme_set(theme_bw())

AllDataSummarized.forplot = temp.summary %>%
  group_by(site,month) %>%
  summarize(., site_month = mean(mean))

p <- ggplot(
  AllDataSummarized.forplot,
  aes(month, site_month, group = site, color = factor(site))
  ) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Month", y = "Temperature") +
  theme(legend.position = "top")

anim_save("bymonth.gif", p + transition_reveal(month))

```

4) GENERAL PLOT---hourly heat map for representative sites: ULAQ1S (73%forested) and PUCC2S (83.8%urban watershed)
Follows the code developed by  John MacKintosh: https://www.r-graph-gallery.com/283-the-hourly-heatmap.html 
```{r setup, include=FALSE}

library(ggplot2)
library(dplyr) 
library(viridis)
library(Interpol.T) 
library(lubridate) 
library(ggExtra)
library(tidyr) 

#first I need to create hourly data for all sites:
file.names <- dir(getwd(), pattern =".csv") #a vector with all file names in the inout folder
temp.summary<-data.frame(matrix(ncol = 9, nrow = 0)) #empty dataframe that will contain the integrated files into one single dataframe
colnames(temp.summary) <- c("year","month","day","hour","min","mean","max","diel","site") #name for the eight columns

for(i in 1:length(file.names)){  
              
        file <- read.table(file.names[i],header=TRUE, sep=",", stringsAsFactors=TRUE)
        data2 <- select(file,c(sitecode,dateEST,TimeEST,depth_mm,temp_degc))
        data2$dateEST <- as.Date(data2$dateEST, format = "%d %B %y")
        data2$dateEST <- as.character(data2$dateEST) ; data2$TimeEST <- as.character(data2$TimeEST)
        data2$datetime <- paste(data2$dateEST, data2$TimeEST) ; data2$datetime <- ymd_hm(data2$datetime)
        temp_data<-select(data2,c(sitecode,depth_mm,temp_degc,datetime))

              temp_data <- temp_data %>%
                mutate(year = year(datetime))  %>%
                mutate(month = month(datetime))  %>%
                mutate(day = day(datetime)) %>%
                mutate(hour = hour(datetime))
              temp_data <- temp_data %>%
                filter(year==c("2018"))
              temp_diel_data <- temp_data %>%
                group_by(year,month,day,hour) %>%
                summarise_at("temp_degc", tibble::lst(min, mean, max, diel))

                        temp_diel_data$site <- as.vector(rep(file.names[i], nrow(temp_diel_data)))
                        colnames(temp_diel_data) <- c("year","month","day","hour","min","mean","max","diel","site")
                        temp_diel_data<-as.data.frame(temp_diel_data)
                        temp.summary<-rbind(temp.summary,temp_diel_data)
                            }
                            
write.csv(temp.summary,file = "R:/EcosystemEcologyLab/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/04_product/AllData/AllDataSummarized_byHour.csv") #write the final file into the sensor product folder

AllDataSummarized_byHour.forPlot = temp.summary %>% 
  select(site,day,hour,month,year,mean) %>%
  filter(site %in% c("ULAQ1S.csv", "PUCC2S.csv")) #to clean the datase for the plot

#Now the plot:
yearno <-unique(AllDataSummarized_byHour.forPlot$year)
 
######## Plotting starts here#####################
p <-ggplot(AllDataSummarized_byHour.forPlot,aes(day,hour,fill=mean))+
  geom_tile(color= "white",size=0.1) + 
  scale_fill_viridis(name="Hrly Temps C",option ="C")
p <-p + facet_grid(site~month)
p <-p + scale_y_continuous(trans = "reverse", breaks = unique(AllDataSummarized_byHour.forPlot$hour))
p <-p + scale_x_continuous(breaks =c(1,10,20,31))
p <-p + theme_minimal(base_size = 8)
p <-p + labs(title= paste("Hourly Temps - Year",yearno), x="Day", y="Hour of Day")
p <-p + theme(legend.position = "bottom")+
  theme(plot.title=element_text(size = 14))+
  theme(axis.text.y=element_text(size=6)) +
  theme(strip.background = element_rect(colour="white"))+
  theme(plot.title=element_text(hjust=0))+
  theme(axis.ticks=element_blank())+
  theme(axis.text=element_text(size=7))+
  theme(legend.title=element_text(size=8))+
  theme(legend.text=element_text(size=6))+
  removeGrid()#ggExtra
 
# you will want to expand your plot screen before this bit!
p #awesomeness

``` 

5) HGAM with hourly temp data of 2018
```{r}

library(mgcv)
library(plyr)
library(viridis)
library(dplyr)
library(tidyr)
library(mvtnorm)
library(statmod)

AllDataSummarized_byHour.forGAM = temp.summary %>% #first for one site and see how it goes
  filter(site %in% c("PUCC2S.csv")) 

GAM_data = AllDataSummarized_byHour.forGAM %>%
  group_by(month,day) %>%
  summarize(day_temp = mean(mean, na.rm = TRUE)) %>%
  arrange(month, day) 
  
GAM_data$c.day=row.names(GAM_data) #rownames as cumulative time factor
GAM_data$day<- as.numeric(GAM_data$day) ; GAM_data$c.day<- as.numeric(GAM_data$c.day)

gam1<- gam(day_temp ~ s(day, bs="cc") + s(c.day, bs='cr',k=300), 
           data=GAM_data, method = 'REML', 
           knots = list (day = c(0.5,24.5)))

summary(gam1)
ggplot(GAM_data, aes(x=c.day,y=day_temp))+geom_line()





```



