 ---
title: "CitSci exploratory"
output: html_document
---

1) Package load/retrieval
```{r,warning=F}
#necessary packages: 
pack_list<-list ("ggplot2","openair","dygraphs","xts","dataRetrieval","dplyr","forecast","lubridate", "expss")
for(i in 1:length(pack_list)){
  p <- pack_list[[i]]
  install.packages(p,repos = "https://cran.rstudio.com",dependencies = TRUE)
}
```


2) INDIVIDUAL SITE ANALISYS
```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)  
library(lubridate)
library(xts)
library(tidyverse)
library(dygraphs)
library(forecast)
library(timeSeries)
library(glue)
library(expss)
detach("package:dplyr", unload=TRUE) : library(dplyr)
options(stringsAsFactors = FALSE)

setwd("~/Library/Mobile Documents/com~apple~CloudDocs/files")
setwd("C:/Users/admin/Dropbox/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/01_input")

#######################################Individial site analysis. Input specs first and check output_lists for results
data <- read.csv("PUWC1S.csv")
data2 <- select(data,c(sitecode,dateEST,TimeEST,depth_mm,temp_degc))
data2$dateEST <- as.Date(data2$dateEST, format = "%d %B %y") #output is YYYY-mm-dd
data2$dateEST <- as.character(data2$dateEST) ; data2$TimeEST <- as.character(data2$TimeEST)
data2$datetime <- paste(data2$dateEST, data2$TimeEST) ; data2$datetime <- ymd_hm(data2$datetime)
temp_data<-select(data2,c(sitecode,depth_mm,temp_degc,datetime))

#create a month variable
temp_data <- temp_data %>%
  mutate(year = year(datetime))  %>%
  mutate(month = month(datetime))  %>%
  mutate(day = day(datetime))
 
temp_summer_data <- temp_data %>%
  filter(month==c("6","7","8","9")) #this dataset will be kept as is    #######    temp_summer_data

diel <- function(x){
  max(x) - min(x)} #to get diel change in T per day

temp_summer_par <- temp_summer_data %>%
  group_by(year,month,day) %>%
  summarise_at("temp_degc", tibble::lst(min, mean, max, diel)) # Vars are Auto named with `tibble::lst()`

```

2) LOOPED SITE MEAN ANALISYS---run October 2019, files saved as products of EnviroDIY
Code selects each file, subset by months 6-9, and summarize by 4 paramaters (min,max,mean, and diel range)
```{r setup, include=FALSE}
### iterate this for each file ###

diel <- function(x){
  max(x) - min(x)} 

file.names <- dir(getwd(), pattern =".csv")
    temp.summary<-data.frame(matrix(ncol = 8, nrow = 1)) 
    colnames(temp.summary) <- c("year","month","day","min","mean","max","diel","site")
            
    for(i in 1:length(file.names)){  
              
file <- read.table(file.names[i],header=TRUE, sep=",", stringsAsFactors=TRUE)
data2 <- select(file,c(sitecode,dateEST,TimeEST,depth_mm,temp_degc))
data2$dateEST <- as.Date(data2$dateEST, format = "%d %B %y")
data2$dateEST <- as.character(data2$dateEST) ; data2$TimeEST <- as.character(data2$TimeEST)
data2$datetime <- paste(data2$dateEST, data2$TimeEST) ; data2$datetime <- ymd_hm(data2$datetime)
temp_data<-select(data2,c(sitecode,depth_mm,temp_degc,datetime))

temp_data <- temp_data %>%
  mutate(year = year(datetime))  %>%
  mutate(month = month(datetime))  %>%
  mutate(day = day(datetime))
temp_summer_data <- temp_data %>%
  filter(month==c("6","7","8","9")) 
temp_summer_par <- temp_summer_data %>%
  group_by(year,month,day) %>%
  summarise_at("depth_mm", tibble::lst(min, mean, max, diel))

temp_summer_par$site <- as.vector(rep(file.names[i], nrow(temp_summer_par)))
colnames(temp_summer_par) <- c("year","month","day","min","mean","max","diel","site")
temp.summary<-rbind(temp.summary,temp_summer_par)
    }
    
write.csv(temp.summary,file = "R:/EcosystemEcologyLab/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/04_product/DailyDataDepth.csv")
```    
    
3) LOOPED SITE HOURS FOR THERMAL STRESS ANALISYS--run November 4th, 2019, file saved as "Temp_criterias.csv" in product folder
```{r}    
setwd("R:/EcosystemEcologyLab/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/03_incremental/01_CitSci_temperature")
criteria=read.csv("Tcriterias.csv")
criteria_temp=select(criteria,c(dateEST,TimeEST,tempCWF,tempWWF,tempTSF))
criteria_temp$dateEST=as.Date(criteria_temp$dateEST, format = "%d-%B-%y")
criteria_temp$dateEST=as.character(criteria_temp$dateEST) ; criteria_temp$TimeEST=as.character(criteria_temp$TimeEST)
criteria_temp$datetime=paste(criteria_temp$dateEST, criteria_temp$TimeEST) ; criteria_temp$datetime=ymd_hm(criteria_temp$datetime)

temp_data<-select(criteria_temp,c(tempCWF,tempWWF,tempTSF,datetime))
criteria=temp_data %>%
  mutate(year = year(datetime))  %>%
  mutate(month = month(datetime))  %>%
  mutate(day = day(datetime)) %>%
  mutate(hour = hour(datetime))
criteria$datetime=paste(criteria$year, 
                          criteria$month, 
                            criteria$day, 
                              criteria$hour)
#this contains criteria for temperatures, time, and date splitted by year/month/day

#Now loop each file similary to section (2)
setwd("R:/EcosystemEcologyLab/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/01_input")
    
file.names=dir(getwd(), pattern =".csv")
    temp.summary<-data.frame(matrix(ncol = 9, nrow = 1)) 
    colnames(temp.summary)=c("tempCWF","tempWWF","tempTSF","year","month","day","hour","mean_Temp","site")
            
    for(i in 1:length(file.names)){  
              
      file=read.table(file.names[i],header=TRUE, sep=",", stringsAsFactors=TRUE)
      data2=select(file,c(sitecode,dateEST,TimeEST,depth_mm,temp_degc))
      data2$dateEST=as.Date(data2$dateEST, format = "%d %B %y")
      data2$dateEST=as.character(data2$dateEST) ; data2$TimeEST=as.character(data2$TimeEST)
      data2$datetime=paste(data2$dateEST, data2$TimeEST) ; data2$datetime=ymd_hm(data2$datetime)
      temp_data<-select(data2,c(sitecode,depth_mm,temp_degc,datetime))
        temp_data=temp_data %>%
            mutate(year = year(datetime))  %>%
            mutate(month = month(datetime))  %>%
            mutate(day = day(datetime)) %>%
            mutate(hour = hour(datetime))
        temp_summer_data=temp_data %>%
            filter(month==c("6","7","8","9")) 
        temp_summer_par=temp_summer_data %>%
          group_by(year,month,day,hour) %>%
          summarise_at("temp_degc", mean)
        temp_summer_par$datetime=paste(temp_summer_par$year, 
                                        temp_summer_par$month, 
                                          temp_summer_par$day, 
                                          temp_summer_par$hour)

temp_summer_par$site=as.vector(rep(file.names[i], nrow(temp_summer_par)))
combined_dataset=left_join(criteria,temp_summer_par,by="datetime")
combined_dataset=select(combined_dataset,c(tempCWF,tempWWF,tempTSF,year.x,month.x,day.x,hour.x,temp_degc,site))
colnames(combined_dataset)=c("tempCWF","tempWWF","tempTSF","year","month","day","hour","mean_Temp","site")
temp.summary<-rbind(temp.summary,combined_dataset)
    }
write.csv(temp.summary,file = "C:/Users/admin/Dropbox/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/04_product/Temp_criterias.csv")
```

4) Continuation of THERMAL STRESS ANALISYS
```{r}
setwd("R:/EcosystemEcologyLab/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/04_product/DRWIWork")
data <- read.csv("Temp_criterias.csv")

combined_dataset2=data
combined_dataset2 <- na.omit(combined_dataset2)
combined_dataset2$Tcount = count_row_if(gt(combined_dataset2$tempCWF), combined_dataset2$mean_Temp)
combined_dataset2$site=as.factor(combined_dataset2$site) 
combined_dataset2 = combined_dataset2[complete.cases(combined_dataset2), ]
tempCWF_stress = combined_dataset2 %>% 
      filter(.,site %in% c("BCWC10S.csv","PUCC2S.csv","PUWC1S.csv","SHPK6S.csv","ULAQ1S.csv","ULLL2S.csv","ROCK_DS2.csv","HURR_US2.csv","MSPR2S.csv","BCMC3S.csv")) %>% 
      filter(.,year %in% c("2018")) %>% 
      group_by(site,year,month,day) %>%
      summarise(Tcount = sum(Tcount))
write.csv(tempCWF_stress,file = "R:/EcosystemEcologyLab/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/04_product/DRWIWork/tempCWF_stress.csv")

combined_dataset2$Tcount = count_row_if(gt(combined_dataset2$tempWWF), combined_dataset2$mean_Temp)
tempWWF_stress = combined_dataset2 %>% 
      filter(.,site %in% c("BCWC10S.csv","PUCC2S.csv","PUWC1S.csv","SHPK6S.csv","ULAQ1S.csv","ULLL2S.csv","ROCK_DS2.csv","HURR_US2.csv","MSPR2S.csv","BCMC3S.csv")) %>% 
      filter(.,year %in% c("2018")) %>% 
      group_by(site,year,month,day) %>%
      summarise(Tcount = sum(Tcount))
write.csv(tempWWF_stress,file = "R:/EcosystemEcologyLab/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/04_product/DRWIWork/tempWWF_stress.csv")

combined_dataset2$Tcount = count_row_if(gt(combined_dataset2$tempTSF), combined_dataset2$mean_Temp)
tempTSF_stress = combined_dataset2 %>% 
      filter(.,site %in% c("BCWC10S.csv","PUCC2S.csv","PUWC1S.csv","SHPK6S.csv","ULAQ1S.csv","ULLL2S.csv","ROCK_DS2.csv","HURR_US2.csv","MSPR2S.csv","BCMC3S.csv")) %>% 
      filter(.,year %in% c("2018")) %>% 
      group_by(site,year,month,day) %>%
      summarise(Tcount = sum(Tcount))
write.csv(tempTSF_stress,file = "R:/EcosystemEcologyLab/DRWIdataDirectory/DataInDevelopment/CitSciDeployments/EnviroDIY/sensors/04_product/DRWIWork/tempTSF_stress.csv")


```    



